Experience shapes individual foraging specialization and success in a virtual predator-prey system

### 2023-06-20 manuscript/Chapter2-manuscript.Rmd
minor updates to answer JM comments .rmd

All models were fitted under a Bayesian framework in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@Burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@Gabry.Cesnovar2021] as the back-end for estimation (cmdstan installation version 2.28.2). The models were compiled on Cedar (Operating system: CentOS Linux 7), a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar). Each required 64GB of RAM with 48 cores to compile within 5 days.

@@

We tested how prey influenced the development of predator expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e. number of prey consumed) and the predators' cumulative experience (i.e. number of matches played before the current match). We parametrized the models following the method of @Pedersen.etal2019. The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID. In this model, we assume that predators have the same development of expertise, with the model estimating a trend for the average individual (i.e. global smoother). For the second model, we included varying individual smoothers for the predator ID. Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g. predator 1 has a steeper curve than predator 2). This enabled us to test whether predators differed in the development of their expertise. In the third model, we kept the individual smoothers for the predators, but removed the global smoother. This model assumes that predators do not share a common relationship between success and experience. The fourth and fifth models were a reproduction of the second and third models respectively, where we included the prey speed to assess its effect on the relationship between success and experience. We included the standardized match duration and prey rank as covariates in all five models.

@@

We tested whether foraging behaviour changed with experience, and compared the success of predators along the flexible-specialist hunter continuum, using a multivariate double-hierarchical generalized linear model (MDHGLM) [@Lee.Nelder2006;@Cleasby.etal2015;@Mitchell.etal2016a]. For each experience level, we fitted the model on three traits: the predator's speed ($Y_{1}$) and the prey's speed ($Y_{2}$) using a Gaussian error distribution, and the predator's hunting success ($Y_{3}$) using a beta-binomial error distribution. Fixed and random effect structure were trait-specific and the same for all experience levels. The predator and the prey speed included a mean and a dispersion part. For both traits, we controlled for the standardized prey rank on both the mean and the dispersion parts of the model. We included game environment, predator avatar, and individual predator identity random effects on the mean part of the model. The dispersion part included individual predator identity random effects exclusively. For hunting success, we modeled the mean part only and controlled for the standardized prey rank and game duration, and included an individual identity random effect.

The model simultaneously estimated three individual components for each trait across experience levels. The first component, estimated on the mean part of the model, is among individual variation in mean speed and success. This allows to test for individual differences in tactic use (i.e. along the slow-fast continuum), in the mean speed of the prey encountered, and in mean prey consumption. The second component, estimated on the dispersion part of the model by modeling the residuals, is among individual variation in intra-individual variation (IIV). This allows to test if predators do not experience the same degree of prey heterogeneity and if they differ in individual specialization. A general intercept is also fitted on both the mean and dispersion part of the model, providing population-level estimates of the aforementioned parameters. The third component consists of pairwise comparisons between the individual identity standard deviations on the mean and dispersion part of the model among traits, among experience, and across traits and experience. This results in a 15x15 (co)variance matrix, allowing to describe the structure of the predator-prey trait interactions as well as the relationship between specialization and success across experience at the individual-level.

@@

where $j$ is the index of the three experience levels (i.e. novice, intermediate, advanced). The $d$ subscript indicates that a term is from the dispersion part of the model. The $b$ terms are vectors of fixed effects with their design matrices $X$. The terms $en$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their design matrices $M$, $N$, and $Z$. For all experience levels of $Y_{1}$ and $Y_{2}$, we assumed that the game environment and avatar random effects followed a Gaussian distribution with estimated standard deviation ($en_{12j} \sim N(0,\ I\ \sigma_{en_{12j}})$ and $av_{12j} \sim N(0,\ I\ \sigma_{av_{12j}})$). $I$ is the identity matrix, with $I\sigma$ indicating that the random effects are independently and identically distributed. In addition, we assumed that the residuals follow a Gaussian distribution with observation-specific standard deviations $e_{12j} \sim N(0,\ Diag\{\sigma_{e_{12j}}\})$. All the individual identity random effects (mean and dispersion) across all response variables and experience levels are indexed in vector $\alpha$. The $\alpha$ vector follows a multivariate Gaussian distribution $\alpha \sim MVN(0,\ \Sigma)$ where $\Sigma$ is a 15x15 (co)variance matrix of the individual identity standard deviations $\sigma_{id_{ij}}$ across all response variables and experience levels.

We used weakly informative Gaussian priors for the prey rank ($N(0,\ 1)$: eqn. 1.1 to 2.2) and game duration ($N(0.5,\ 0.5)$: eqn. 3). Based on previous results showing that the mean predator speed revolves around 3 m/s [@FraserFranco.etal2022], we applied a weakly informative Gaussian prior on the intercept for predator and prey speed ($N(3,\ 0.5)$: eqn. 1.1 and 2.1). Similarly, we used a weakly informative Gaussian prior on the intercept for the dispersion part ($N(0,\ 1)$: eqn 1.2 and 2.2). For hunting success, we applied a positive Gaussian prior on the precision parameter ($N(2,\ 1)$: eqn. 3) and a negative Gaussian prior ($N(-4,\ 1)$: eqn 3) on the intercept because hunting success should be close to 0 at the lowest game duration. For every predicted variable, we applied a weakly informative half-Gaussian prior ($N(0,\ 1)$) on the random effects for the mean part and the dispersion part of the model. We used a regularizing $LKJ(2)$ prior on the correlation matrix.

@@

Our results suggest that predator expertise is honed through extensive practice. The predator population increased and stabilized its success with experience in a diminishing returns fashion as it is typically found in empirical studies of expertise [reviewed in @Dukas2019]. The prey were important in mediating this pattern, probably because they increased their speed as predators gained experience. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@FraserFranco.etal2022], and other studies have found that as well [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. This resulted in discernable differences in the relationship between success and experience among predators, implying that the development of expertise depends on an individual's capacity to adjust to difficulty. Hunting faster prey requires costly and specialized cognitive abilities and coordination in predators [@Kelley.Magurran2011], and that could explain why such prey were more difficult to capture. Thus, predators that couldn't properly hunt at high speeds or develop counterstrategies for faster prey were likely at a disadvantage.

However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predators, suggesting that other antipredator tactics were potentially involved. For example, if the prey used camouflage or hiding to avoid being detected [@Bowen.etal2002;@Carey.Wahl2011;@Nomura.etal2011], then predators may have found them only if they learned how to exploit the visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@Ehlinger1989;@Hughes.etal1992;@Warburton2003;@Szopa-Comley.etal2020]. In addition, longer time intervals between hunting events may delay or even impede learning due to forgetting important information [@Endler1991;@Wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Accounting for such time lags may reveal important insights on the outcome of predator-prey interactions.

@@

The predator population maintained its speed and degree of specialization throughout the study period, and individuals differed only slightly in their average speeds. However, differences in specialization among individuals were important and increased slightly with experience. Predators hunting at higher speeds specialized in this tactic and tended to encounter groups of prey with similar speeds, while those hunting at slower speeds were flexible and tended to encounter varying groups of prey. On one hand, this pattern could arise if some predators adjusted their strategy to their prey, because fast-paced hunting is a specialized tactic suited for prey that use rapid evasive movements [@Endler1991;@Bro-Jorgensen2013;@Wilson.etal2018], while a slower and flexible tactic can be useful to minimize the consequences of uncertainty when prey escape unpredictably [@Szopa-Comley.Ioannou2022]. However, we observed that 44% of the predator population showed close to no change in specialization with experience. Since these individuals were already specialized as novices (Fraser Franco, personal observation), there would be no point in them changing their tactic with experience if they were already successful, or there could have been costs to switch their strategy if they more often encountered faster prey. This could explain why the behavioural interactions between the predator and the prey remained stable across the predator's experience. Yet, because prey can also learn how to avoid predation [@kelley2003;@turner2006;@mccomb2011], we believe that the predator-prey phenotype matching more likely emerged from reciprocal adjustments in speed by predators and prey as they interacted [@Kishida.etal2006;@Kishida.etal2009;@Edgell.Rochette2009a;@McGhee.etal2013a]. Indeed, 44% of the predator population altered its behaviour with experience to different degrees, and 12% even switched between flexibility and specialization. Hence, if the prey also learned through repeated interactions with the predators, it is possible that experience contributed in stabilizing the system as both were adusting to each other, similar to Red Queen dynamics [@brockhurst2014]. Our results appear to support this since many individuals were shifting their behavioural niche while the population niche remained stable across experience.

Classical theory predicts that individual foraging specialization should maximize prey consumption by optimizing foraging efficiency [@stephens1986], but other work suggests that flexibility and specialization could both be adaptive as resources fluctuate [@Woo.etal2008;@Phillips.etal2017]. Our results support empirical findings indicating that flexible and specialist foragers obtained similar success [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015;@Phillips.etal2017;@Courbin.etal2018], although flexible foragers were slightly less successful. Both strategies appeared to emerge in part from individuals learning how to hunt their prey, resulting in a general increase in success in the population. However, there were still considerable differences in success among individuals through time, suggesting that some predators were limited in their capacity to match their tactic to their prey or to increase their success. Specialist foragers were faster, and thus, probably better equipped to hunt the more difficult faster prey in *DBD*. However, if the prey responded to fast predators by also being faster, then hunting at high speeds resulted in more difficult encounters for the predator, thereby nullifying the benefits of using this tactic (Figure 4). Thus, specializing probably compensated for the difficulty of hunting prey at high speeds by helping predators to better predict the location and movement of their prey. On the other hand, flexible foragers encountered variable prey with slower speeds. Yet, because the prey increased their speed with experience, the benefits of being able to hunt multiple prey types for flexible hunters may have come at the cost of not being adept at capturing faster prey [@Pintor.etal2014]. Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@Dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey. As the benefits of each hunting style changed under different scenarios, the combination of predatory styles in our virtual system reflects how resource fluctuations can maintain fitness equilibrium within populations, resulting in behavioural diversity in predator-prey systems [@Tinker.etal2008;@Woo.etal2008;@Mougi.Kishida2009].

@@

We warmly thank Behaviour Interactive Inc. for our collaboration which helps in advancing predator-prey science. We thank Julien Céré who managed the communication between us and Behaviour Interactive Inc., and Marine Dupuy who helped us prepare the data used in this study. We thank Alastair J. Wilson for insightful comments on the results. We also thank all members of Pierre-Olivier Montiglio's laboratory who provided useful feedback on earlier versions on this work. This work was supported by an NSERC postgraduate doctoral scholarship (569716-2022).

@@

The authors declare no conflict of interest

@@

**Figure 1.** Median posterior predictions of the development of predator hunting expertise. The predators' hunting success (i.e. the probability of consuming the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Panels A and C show the development of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values. For panel A, the right-side vertical dashed lines shows the highest predicted success. For panel C, the right-side dashed line represents the point on the curve where success was optimized, which we calculated using the finite differences method to obtain the first derivative of the predicted values. Panels B and D show among individual differences in the development of expertise, with each curve representing an individual predator. (A-B) GAMM where we control for the prey rank (C-D) GAMM where we control for the prey rank and the speed of the prey group.

@@

**Figure 3.** Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility. The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters. Both figure panels are ordered by ascending degree of increase in either specialization or flexibility. We built the figure by first subtracting the estimated standard deviation of all individuals as novices with their standard deviation as advanced hunters, and then selected those with the greatest difference using the 25% and 75% quantiles. Individuals on panel A have an increase in specialization equal or greater than 0.2 standard deviations, while those on panel B have an increase in flexibility equal or greater than 0.2 standard deviations.

**Figure 4.** Correlation matrices of the MDHGLM relating the mean and IIV in speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Larger dark blue circles indicate stronger positive correlations (positive values on the colour legend), while larger dark red circles indicate stronger negative correlations (negative values on the colour legend). A) Correlations when predators were novice hunters. B) Correlations when predators were advanced hunters. C) Posterior median differences and HPD intervals comparing the trait correlations predicted by the MDHGLM between the advanced and novice levels. The difference is displayed on the y axis and the correlated traits being compared are displayed on the x axis.Positive values indicate an increase in correlation strength from novice to the advanced stage, while negative values indicate stronger correlations when predators were novice.

>>>>

All models were fitted under a Bayesian framework in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@Burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@Gabry.Cesnovar2021] as the back-end for estimation (cmdstan installation version 2.28.2). The models were run on Cedar (Operating system: CentOS Linux 7), a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar). Each required 64GB of RAM with 48 cores to compile within 5 days.

@@

We tested how prey influenced the development of predator expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e. number of prey consumed) and the predators' cumulative experience (i.e. number of matches played before the current match). We parametrized the models following the method of @Pedersen.etal2019. The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID. In this model, we assume that predators have the same development of expertise, with the model estimating a trend for the average individual (i.e. global smoother). For the second model, we included varying individual smoothers for the predator ID. Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g. predator 1 has a steeper curve than predator 2). This enabled us to test whether predators differed in the development of their expertise. In the third model, we kept the individual smoothers for the predators, but removed the global smoother. This model assumes that predators do not share a common relationship between success and experience. The fourth and fifth models were reproductions of the second and third models respectively, where we included the prey speed to assess its effect on the relationship between success and experience. We included the standardized match duration and prey rank as covariates in all five models.

@@

We tested whether foraging behaviour changed with experience, and compared the success of predators along the flexible-specialist hunter continuum, using a multivariate double-hierarchical generalized linear model (MDHGLM) [@Lee.Nelder2006;@Cleasby.etal2015;@Mitchell.etal2016a]. For each experience level, we fitted the model on three traits: the predator's speed ($Y_{1}$) and the prey's speed ($Y_{2}$) using a Gaussian error distribution, and the predator's hunting success ($Y_{3}$) using a beta-binomial error distribution. Fixed and random effect structure were trait-specific and the same for all experience levels. The predator and the prey speed included a mean and a dispersion part (eqn 1.1, 1.2 and 2.1, 2.2). For both traits, we controlled for the standardized prey rank on both the mean and the dispersion parts of the model. We included game environment, predator avatar, and individual predator identity random effects on the mean part of the model. The dispersion part included individual predator identity random effects exclusively. For hunting success, we modeled the mean part only (eqn 3) and controlled for the standardized prey rank and game duration, and included an individual identity random effect.

The model simultaneously estimated five among-individual variance components with their covariances for each experience level, resulting in a 15x15 (co)variance matrix. For each experience level, the model estimated among individual variance in mean predator speed, mean prey speed, and mean hunting success. This allows to test for individual differences in tactic use (i.e. along the slow-fast continuum), in the mean speed of the prey encountered, and in mean prey consumption. The model also estimated among individual variance in intra-individual variation (IIV) for both prey and predator speed. This allows to test if predators do not experience the same degree of prey heterogeneity and if they differ in individual specialization. Finally, the model estimated all pairwise covariances at the individual level among traits, among experience, and across traits and experience. The resulting 15x15 (co)variance matrix thus provides the structure of the predator-prey trait interactions as well as the relationship between specialization and success across experience at the individual-level.

@@

where $j$ is the index of the three experience levels (i.e. novice, intermediate, advanced). The $d$ subscript indicates that a term is from the dispersion part of the model. The $b$ terms are vectors of fixed effects with their design matrices $X$ specific to each trait and experience level. The terms $en$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their design matrices $M$, $N$, and $Z$ specific to each trait and experience level. For all experience levels of $Y_{1}$ and $Y_{2}$, we assumed that the game environment and avatar random effects followed a Gaussian distribution with estimated standard deviation ($en_{12j} \sim N(0,\ I\ \sigma_{en_{12j}}^{2})$ and $av_{12j} \sim N(0,\ I\ \sigma_{av_{12j}}^{2})$). $I$ is the identity matrix, with $I\sigma^{2}$ indicating that the random effects are independently and identically distributed. In addition, we assumed that the residuals follow a Gaussian distribution with observation-specific variance $e_{12j} \sim N(0,\ Diag\{\sigma_{e_{12j}}^{2}\})$. The 15 individual identity random effects $id_{ij}$ across all response variables and experience levels (five for each of the three experience levels) are indexed in vector $\alpha$. The $\alpha$ vector follows a multivariate Gaussian distribution $\alpha \sim MVN(0,\ \Sigma)$ where $\Sigma$ is a 15x15 (co)variance matrix of among individual variation across all response variables and experience levels, with $\sigma_{id_{ij}}^{2}$ on the diagonal and covariances off-diagonal.

We used weakly informative Gaussian priors for the prey rank ($N(0,\ 1^{2})$: eqn. 1.1 to 2.2) and game duration ($N(0.5,\ 0.5^{2})$: eqn. 3). Based on previous results showing that the mean predator speed revolves around 3 m/s [@FraserFranco.etal2022], we applied a weakly informative Gaussian prior on the intercept for predator and prey speed ($N(3,\ 0.5^{2})$: eqn. 1.1 and 2.1). Similarly, we used a weakly informative Gaussian prior on the intercept for the dispersion part ($N(0,\ 1^{2})$: eqn 1.2 and 2.2). For hunting success, we applied a positive Gaussian prior on the precision parameter ($N(2,\ 1^{2})$: eqn. 3) and a negative Gaussian prior ($N(-4,\ 1^{2})$: eqn 3) on the intercept because hunting success should be close to 0 at the lowest game duration. For every predicted variable, we applied a weakly informative half-Gaussian prior ($N(0,\ 1^{2})$) on the random effects for the mean part and the dispersion part of the model. We used a regularizing $LKJ(2)$ prior on the correlation matrix.

@@

Our results suggest that predator expertise is honed through extensive practice. The predator population increased and stabilized its success with experience in a diminishing returns fashion as it is typically found in empirical studies of expertise [reviewed in @Dukas2019]. The prey were important in mediating this pattern, probably because they increased their speed as predators gained experience. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@FraserFranco.etal2022], and other studies have found that as well [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. This resulted in discernable differences in the relationship between success and experience among predators, implying that the development of expertise depends on an individual's capacity to adjust to difficulty. Hunting faster prey requires costly and specialized cognitive abilities and coordination in predators [@Kelley.Magurran2011], and that could explain why such prey were more difficult to capture. Thus, predators that couldn't properly hunt at high speeds or develop counter-strategies for faster prey were likely at a disadvantage.

However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predators, suggesting that other antipredator tactics were potentially involved. For example, if the prey used camouflage or hiding to avoid being detected [@Bowen.etal2002;@Carey.Wahl2011;@Nomura.etal2011], then predators may have found them only if they learned how to exploit the visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@Ehlinger1989;@Hughes.etal1992;@Warburton2003;@Szopa-Comley.etal2020]. In addition, longer time intervals between hunting events may delay or even impede learning due to forgetting important information [@Endler1991;@Wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.

@@

The predator population maintained its speed and degree of specialization throughout the study period, and individuals differed only slightly in their average speeds. However, differences in specialization among individuals were important and increased slightly with experience. Predators hunting at higher speeds specialized in this tactic and tended to encounter groups of prey with similar speeds, while those hunting at slower speeds were flexible and tended to encounter varying groups of prey. On one hand, this pattern could arise if some predators adjusted their strategy to their prey, because fast-paced hunting is a specialized tactic suited for prey that use rapid evasive movements [@Endler1991;@Bro-Jorgensen2013;@Wilson.etal2018], while a slower and flexible tactic can be useful to minimize the consequences of uncertainty when prey escape unpredictably [@Szopa-Comley.Ioannou2022]. However, we observed that 44% of the predator population showed close to no change in specialization with experience. Since these individuals were already specialized as novices (Fraser Franco, personal observation), there would be no point in them changing their tactic with experience if they were already successful, or there could have been costs to switch their strategy if they more often encountered faster prey. This could explain why the behavioural interactions between the predator and the prey remained stable across the predator's experience. Yet, because prey can also learn how to avoid predation [@kelley2003;@turner2006;@mccomb2011], we believe that the predator-prey phenotype matching more likely emerged from reciprocal adjustments in speed by predators and prey as they interacted [@Kishida.etal2006;@Kishida.etal2009;@Edgell.Rochette2009a;@McGhee.etal2013a]. Indeed, 56% of the predator population altered its behaviour with experience to different degrees, including 12% that even switched between flexibility and specialization. Hence, if the prey also learned through repeated interactions with the predators, it is possible that experience contributed in stabilizing the system as both were adjusting to each other, similar to Red Queen dynamics [@brockhurst2014]. Our results appear to support this since many individuals were shifting their behavioural niche while the population niche remained stable across experience.

Classical theory predicts that individual foraging specialization should maximize prey consumption by optimizing foraging efficiency [@stephens1986], but other work suggests that flexibility and specialization could both be adaptive as resources fluctuate [@Woo.etal2008;@Phillips.etal2017]. Our results support empirical findings indicating that flexible and specialist foragers obtained similar success [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015;@Phillips.etal2017;@Courbin.etal2018], although flexible foragers were slightly less successful. Both strategies appeared to emerge in part from individuals learning how to hunt their prey, resulting in a general increase in success in the population. However, there were still considerable differences in success among individuals through time, suggesting that some predators were limited in their capacity to match their tactic to their prey or to increase their success. Specialist foragers were faster, and thus, probably better equipped to hunt the more difficult faster prey in *DBD*. However, if the prey responded to fast predators by also being faster, then hunting at high speeds resulted in more difficult encounters for the predator, thereby decreasing the benefits of using this tactic (Figure 4). Thus, specializing probably compensated for the difficulty of hunting prey at high speeds by helping predators to better predict the location and movement of their prey. On the other hand, flexible foragers encountered variable prey with slower speeds. Yet, because the prey increased their speed with experience, the benefits of being able to hunt multiple prey types for flexible hunters may have come at the cost of not being adept at capturing faster prey [@Pintor.etal2014]. Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@Dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey. As the benefits of each hunting style changed under different scenarios, the combination of predatory styles in our virtual system reflects how resource fluctuations can maintain fitness equilibrium within populations, resulting in behavioural diversity in predator-prey systems [@Tinker.etal2008;@Woo.etal2008;@Mougi.Kishida2009].

@@

We warmly thank Behaviour Interactive Inc. for our collaboration which helps in advancing predator-prey science. We thank Julien Céré who managed the communication between us and Behaviour Interactive Inc., and Marine Dupuy who helped us prepare the data used in this study. We thank Alastair J. Wilson for insightful comments on the results. We also thank all members of Pierre-Olivier Montiglio's laboratory who provided useful feedback on earlier versions of this work. This work was supported by an NSERC postgraduate doctoral scholarship (569716-2022).

@@

The authors declare no conflict of interest.

@@

**Figure 1.** Median posterior predictions of the development of predator hunting expertise. The predators' hunting success (i.e. the probability of consuming the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Panels A and C show the development of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values. For panel A, the right-side vertical dashed line shows the highest predicted success. For panel C, the right-side dashed line represents the point on the curve where success was optimized, which we calculated using the finite differences method to obtain the first derivative of the predicted values. Panels B and D show among individual differences in the development of expertise, with each curve representing an individual predator. (A-B) GAMM where we control for the prey rank (C-D) GAMM where we control for the prey rank and the speed of the prey group.

@@

**Figure 3.** Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility. The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters. Both figure panels are ordered by ascending degree of increase in either specialization or flexibility. We built the figure by first subtracting the estimated standard deviation of all individuals as novices with their standard deviation as advanced hunters, and then selected those with the greatest difference using the 25% and 75% quantiles. Individuals on panel A have an increase in specialization equal or greater than 0.2 standard deviations, while those on panel B have an increase in flexibility equal to or greater than 0.2 standard deviations.

**Figure 4.** Correlation matrices of the MDHGLM relating the mean and IIV in speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Larger dark blue circles indicate stronger positive correlations (positive values on the colour legend), while larger dark red circles indicate stronger negative correlations (negative values on the colour legend). A) Correlations when predators were novice hunters. B) Correlations when predators were advanced hunters. C) Posterior median differences and HPD intervals comparing the trait correlations predicted by the MDHGLM between the advanced and novice levels. The difference is displayed on the y axis and the correlated traits being compared are displayed on the x axis. Positive values indicate an increase in correlation strength from novice to the advanced stage, while negative values indicate stronger correlations when predators were novice.
